---
title: "Fiches Belges"
author: "Remco Niggebrugge"
date: "01/29/2015"
output: html_document
---

This RMarkdown file will create Fiches Belges statistics for the EJN website. Some pre-processing of the raw server-log files is required. Starting from these raw files, the following steps need to be taken prior to running this processing code.

* include only lines matching " /ejn/ " or ".aspx"
* exclude all files matching " 443 " or "Error" or "LostSession" or "spider"
* from the remaining lines, only keep columns 1,5 and 6 (date, uri, query)
* 

####Preprocessing

Step 1: grepping only webpages, per day.

Assuming shell is on folder "0_raw-files"

<pre>
for i in *.log   
do
cat $i | grep -i " /ejn/ \|.aspx" | grep -vi "spider\| 443 \|ErrorPage\|LostSession" > ../1_processed-days/$i
done

for i in *.log
do
awk '{print $1,$5,$6}' ../1_processed-days/$i > ../2_stripped-days/$i
done

cat ../2_stripped-days/u_ex141* > ../4_quarters/2014Q4.txt

for i in 10 11 12
do
cat ../2_stripped-days/u_ex14$i* > ../3_stripped-months/2014-$i.txt
done
</pre>

####Loading

Libraries:

```{r}
library(dplyr)
library(tidyr)
```

Processing Q4 2014, first load table:

```{r load-data, cache=TRUE}
d <- read.table("../4_quarters/2014Q4.txt", sep="", stringsAsFactors = FALSE)
#d <- read.table("../3_stripped-months/2014-11.txt", sep="", stringsAsFactors = FALSE)
```

Reshape data and only keep Fiches Belges Results pages:

```{r filter-and-reshape, cache=TRUE}
d <- d[,2:3]
names(d) <- c("u","q")
d <- d[grep("FichesBelgesResult.aspx",d$u),]
d <- d[grep("measure=[0-9]*&country=[0-9A-Z]*&other=", d$q),]
d$q <- gsub("measure=","",d$q)
d$q <- gsub("&country="," ",d$q)
d$q <- gsub("&other="," ",d$q)
```

```{r separate-gathering, cache=TRUE}
d <- separate(d, q, into=c("measure","country", "other"), sep=" " )
d <- d[,2:4]
d <- gather(d,temp, country,-measure)
d <- d[,c(1,3)]
d <- d[grep("^[0-9A-Z]", d$country),]
d$c <- as.numeric(d$country)
```

Fixing countries codes:

```{r fixing-country-code, cache=TRUE}
source("../support-files/country-codes.R")
for(i in 1:dim(countries)[[1]]){
        code = countries$code[i]
        short = countries$short[i]
        d$c[d$country==short] <- code
        }

## remove entries did still did not match anything
d <- d[!is.na(d$c),]
d <- merge(d, countries, by.x="c", by.y="code")
d <- d[,c(2,4)]
names(d) <- c("measure","country")
```

We are finally in a position to group the results, we are not interested in individual consultations, but in numbers per measure, per country.

```{r summarize, cache=TRUE}
g     <- group_by(d, measure, country) 
stats <- summarize(g, n = n())
d     <- as.data.frame(stats)

## for plotting I prefer to have 0s values than non-existing entries
for(m in unique(d$measure)){
        for(c in unique(d$country)){
                d <- rbind(d, c(m, c, 0))
        }
}

d$n   <- as.numeric(d$n)

g     <- group_by(d, measure,  country)
stats <- summarize(g, n = sum(n))
d     <- as.data.frame(stats)
```

Creating a table with countries as column names and measures as row names. Then adding totals per column/row.

```{r spreading-table, cache=TRUE}
dd <- spread(d, country, n)
rownames(dd) <- dd[,1]
dd <- select(dd,c(-1))
names(dd) <- countries$short

dd <- cbind(dd, apply(dd,1,sum))
dd <- rbind(dd, apply(dd,2,sum))
names(dd)[dim(dd)[2]] <- "+"
rownames(dd)[dim(dd)[1]] <- "+"

dd

```
